### 2020-04-22 MSCI718 2020W Project v3.Rmd
started report

---
title: "MSCI 718 Mini-Project"
author: "Matthieu Court, Ashar Kandathil, Tyler Lam, Kevin Lee"
date: "April 22, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Instructions: Please complete this assignment and submit online to the LEARN Dropbox as a PDF; the PDF will be considered complete for marking. Also upload any source files, for example, any .R or .Rmd files that you used.*

The final project deliverable must include a **PDF report (max 6 pages, references not included)** and any source code for your project.

**NOTE:** the emphasis **must** be on the application, interpretation, and communication of the statistical method. As part of this, you are **not allowed** to do black-box methods like neural nets. As you've seen, this course is on working data in a transparent and interpretable manner, not on big data analytics.

#Data
Due to the current state of the world from the COVID-19 pandemic, this report focused on using exploratory data analysis, time series forecasting, and data visualization to forecast the number of incidents, recoveries and deaths on a per country basis.

The Global Coronavirus Dataset was sourced from Kaggle and used for this analysis. 
There were three appropriate datasets that were sourced from Kaggle:
1. Global Coronavirus Dataset

We found a dataset that contains information on every country’s number of confirmed cases, deaths, and recoveries since January 1, 2020.

2. Countries of the World Fact Sheet Dataset
3. UN Country Statistics Dataset

We also found 2 other datasets that contain national information on each country, ranging from population to GDP per capita. By combining these 2 datasets, we want to create an interactive dashboard that forecasts the number of cases, recoveries, and deaths per country based on potential correlation with other factors.

Datasets:

#Analysis

#Conclusion